R Tools for Obspack, Footprints and Receptors (rtorf)

### 2024-04-24 examples/README_nc.Rmd
documentation and details ins NEWS.md

categories <- c(
  "aircraft-pfp",
  "aircraft-insitu",
  "surface-insitu",
  "aircore",
  "surface-pfp",
  "tower-insitu",
  "shipboard-insitu",
  "flask"
)
obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs)

@@

```{r, eval = T, include = F}

categories <- c(
  "aircraft-pfp",
  "aircraft-insitu",
  "surface-insitu",
  "aircore",
  "surface-pfp",
  "tower-insitu",
  "shipboard-insitu",
  "flask"
)
obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/txt/"
index <- obs_summary(obs = obs, verbose = T)

There are 362 files in the ObsPack directory.
The printed information also shows the total at the bottom,
as the sum of the individual file by sector. This
is to ensure that the sum of files is equal to the
total number of files found, shown at the top.
furthermore, the printed information
also shows that there are 136 files with the
`agl` explicitly mentioned in the name of the file.

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".
```{r}
dft <- obs_table(index = index,
                 categories = "tower-insitu",
                 verbose = FALSE)
dft[site_code == "SCT", ]$timeUTC |> 
  range()
```

@@

```{r}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

@@

@@

```{r}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14:15

@@

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

@@

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
```

@@

```{r}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

>>>>

```{r, eval = T}

cate = c("aircraft-pfp",
         "aircraft-insitu",
         "aircraft-flask",
         "surface-insitu",
         "surface-flask", 
         "surface-pfp",   
         "tower-insitu",  
         "aircore",       
         "shipboard-insitu",
         "shipboard-flask")

obs <- "Z:/torf/obspack_ch4_1_GLOBALVIEWplus_v5.1_2023-03-08/data/nc/"
index <- obs_summary(obs = obs, categories = cate)

@@

There are `r nrow(index)` files in the ObsPack directory.
The printed information also shows the total at the bottom,
as the sum of the individual file by sector. This
is to ensure that the sum of files is equal to the
total number of files found, shown at the top.
furthermore, the printed information
also shows that there are 109 files with the
`agl` explicitly mentioned in the name of the file.

@@

```{r}
df <- obs_read_nc(index = index,
               categories = "tower-insitu")

@@

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".

```{r}
dft <- obs_table(df = df)
dft[site_name %in% unique(site_name)[sample(1:length(unique(dft$site_code)), 2)]]
```

@@

```{r}
north <- 80
south <- 10
west <- -170
east <- -50
max_altitude <- 8000
evening <- 14

@@

if(any(names(df) %in% "intake_height")) {
  df <- df[intake_height < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
} else {
  df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
}

unique(df[, c("intake_height", "site_elevation", "elevation",
              "dataset_selection_tag",
              "site_name")])

@@

```{r}

if(any(names(df) %in% "intake_height")) {
dfa <- df[,
          max(intake_height),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
} else {
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

names(dfa)[2] <- "max_altitude"
}
dfa
```

@@

```{r}
df2 <- obs_addtime(df)
```

"_st".

```{r}
df2 <- obs_addstime(df2)

```
solar time is not exact, then I need to cut every hour and aggregate observations
by hour. At the end I want one daily observation per day and site

# TODO ...
```{r}
df2 <- df2[hour_st %in% evening]

### 2024-04-24 examples/README_text.Rmd
documentation and details ins NEWS.md

>  This example reads text files but it will be updated to read NetCDF

categories <- c(
  "aircraft-pfp",
  "aircraft-insitu",
  "surface-insitu",
  "aircore",
  "surface-pfp",
  "tower-insitu",
  "shipboard-insitu",
  "flask"
)
obs <- "../../obspack_ch4_1_GLOBALVIEWplus_v4.0_2021-10-14/data/txt"
index <- obs_summary(obs = obs)

@@

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".
```{r}
dft <- obs_table(index = index,
                 categories = "tower-insitu",
                 verbose = FALSE)
dft[site_code == "SCT", ]$timeUTC |> 
  range()
```

@@

```{r}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)

@@

@@

df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
```

@@

```{r}
dfa <- df[,
          max(altitude_final),
          by = site_code] |> unique()

@@

```{r}
df2 <- obs_addtime(df)
df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")
df3 <- obs_addltime(df2)
df3 <- df3[lh %in% evening]
```

>>>>

>  This example reads text files but it will be updated to read text file

@@

@@

```{r}
df <- obs_read(index = index,
               categories = "tower-insitu",
               verbose = FALSE)
df[1:5]

@@

Sometimes we need more information about the site. For instance,
what do the observations start and end. Then, we added 
the function `obs_table`, which calculates statistics
summary of "time" and other numeric variables by file name,
sector, site, altitude and mode. For instance,
the observations in the site "SCT" in
South Carolina, USA, were between "2015-08-19 21:30:00 UTC" and
"2020-12-31 23:30:00 UTC".

```{r}
dft <- obs_table(df = df)
dft[site_name %in% unique(site_name)[1:2]]
# dft[site_code == "SCT", ]$timeUTC |> 
# range()
```

@@

if(any(names(df) %in% "intake_height")) {
  df <- df[intake_height < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
} else {
  df <- df[altitude_final < max_altitude &
           latitude < north &
           latitude > south &
           longitude < east &
           longitude > west]
}

@@

```{r}

df5[,
    max_altitude := max(altitude_final),
    by = site_code]

@@

```{r}
df2 <- obs_addtime(df)
```

```{r}

df2$timeUTC <- cut(x = df2$timeUTC+3600,
                   breaks = "2 hour") |>
  as.character() |>
  as.POSIXct(tz = "UTC")

```{r}
df3 <- obs_addltime(df2)

```{r}
df3 <- df3[lh %in% evening]
